Supplement 1: Testing different distributions

### 2018-09-21 pkg/disttree/inst/draft/supplement/supplement_distributions.Rnw
revised versions of the two supplements
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2288 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 4fd840560f1fff3def8ab5426de9972e78214d72

%\documentclass[a4paper,11pt]{article}
\documentclass[nojss, shortnames]{jss}

@@

\Abstract{In the application discussed in the paper ''Distributional 
Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain''
the underlying distribution is assumed to a be left-censored at zero Gaussian distribution.
In this suppplement the exact same evaluations as in the paper are performed employing
other distribution assumption. In particular, a left-censored at zero logistic distribution 
and a hurdle-model are applied and their results are compared to those presented in the paper.}

\Keywords{parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, GAMLSS}

@@

  Torsten Hothorn\\
  Universit\"at Z\"urich \\
  Institut f\"ur Epidemiologie, Biostatistik und Pr\"avention \\
  Hirschengraben 84\\
  CH-8001 Z\"urich, Switzerland \\
  E-mail: \email{Torsten.Hothorn@R-project.org}\\
  URL: \url{http://user.math.uzh.ch/hothorn/}\\

}

@@

\section{Introduction}
This application study is supplementary material to the paper ''Distributional 
Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain''. 
In Section~3 the presented distributional forest model is employed to obtain 
probabilistic precipitation forecasts in a mountainous region assuming that the 
response variable (the power-transformed daily precipitation sum) follows a 
left-censored Gaussian distribution with censoring point~$0$.

@@

To compare the performance of the distributional forest to other commonly used methods the exact same models as in the paper are evaluated on the same data set and applying the same settings while only the distribution is exchanged. Further details on the models can be found in Section~3.2.. In particular, Table~2 gives an overview over the models and their specifications. As a brief reminder the approaches are summarized below.

@@

\item \emph{Distributional forest:} All predictor variables are
considered for learning a forest where bootstrap
sampling is employed for each tree. Parameters are estimated by
adaptive local likelihood based on forest weights.

@@

The distribution applied in the paper is based on a Gaussian distribution 
for which the probability density function $f$ and the distribution 
function $F$ are defined by
\begin{align}
f_{\textit{Gaussian}}(Y; \mu, \sigma) &= 
     \frac{1}{\sigma} \cdot \phi\left(\frac{Y-\mu}{\sigma}\right) \\
F_{\textit{Gaussian}}(Y; \mu, \sigma) &=
    \Phi\left(\frac{Y-\mu}{\sigma}\right)
\end{align}
where $\phi$ and $\Phi$ are the probability density function and the 
distribution function of the standard normal distribution $\mathcal{N}(0,1)$.

Alternatively one could also choose a distribution with heavier tails 
such as a logistic distribution with corresponding probability density 
function and distribution function

\begin{align}
f_{\textit{logistic}}(Y; \mu, \sigma) &= 
     \frac{\exp{\left(-\frac{Y - \mu}{\sigma}\right)}}
     {\sigma \cdot \left(1 + \exp{\left(-\frac{Y - \mu}{\sigma}\right)}\right)^2} \\
F_{\textit{logistic}}(Y; \mu, \sigma) &=
     \frac{1}{1+ \exp{\left(-\frac{Y - \mu}{\sigma}\right)}}.
\end{align}

Both distributions have to be adapted in our application to capture that
precipitation is non-negative with a point mass at zero (= no rain).
Two common strategies for this are to employ either \emph{left-censoring at zero}
or a \emph{hurdle} (or two-part) approach, combining a binary (zero vs.\ greater) and
a truncated (postive) part \citep[see][for an overview]{Winkelmann:2006}.
\begin{itemize}
\item Censored distribution \citep{Winkelmann:2006}: 
%In general, censoring data usually leads 
%to a point mass at the censoring point which has to be considered in 
%the modeling process. Therefore, censored distributions are applied as 
%they are able to capture this point mass. 
%In precipitation data, observations can not be negative, thus, left-censoring
%it at zero does not change the data. However, there is a naturally given
%point mass at zero (days without rain) making a censored distribution an 
%appropriate choice to model precipitation data.
For a chosen distribution with probability density function $f$ and
distribution function $F$ the log-likelihood function of the corresponding 
censored distribution is
\begin{equation}
\ell_{\textit{cens}}(\mu, \sigma; Y) = 
\begin{cases}
    \log\left\{f(Y; \mu, \sigma) \right\}, & \text{if } Y > 0 \\[0.2cm]
    \log\left\{F(0; \mu, \sigma) \right\}, & \text{if } Y = 0
\end{cases}
\end{equation}
as seen in Equation~2.1.1. of the paper for a Gaussian distribution.
For positive values of $Y$ the censored version of the density function
equals the original non-censored version. For $Y=0$ the censored density
function takes the values of the probability mass lying left of the 
censoring point. Therefore, the parameters $\mu$ and $\sigma$ control
both parts, the point mass and the positive part of the distribution.

\item Hurdle model / Two-part model (\cite{Cragg:1971}, \cite{Mullahy:1986}): 
An additional parameter $\nu$ describing the probability that there is 
any precipitation at all, i.e. $\nu = \mathbb{P}(Y>0)$, is included in the model.
%Then, a second model is fit to the subset of the data containing 
%only positive observations applying a left-truncated at zero 
%distribution to estimate the location parameter $\mu$ and scale 
%parameter $\sigma$ of the underlying latent variable.
The three-parametric log-likelihood function for the full hurdle model 
with parameter vector $\bm{\theta} = (\mu, \sigma, \nu)$ is
\begin{equation} \label{eq:hurdle}
\ell_{\textit{hurdle}}(\mu, \sigma, \nu; Y) = 
\begin{cases}
    \log\left\{\nu \cdot \frac{f(Y; \mu, \sigma)}{1 - F(0; \mu, \sigma)} \right\}, & \text{if } Y > 0 \\[0.2cm]
    \log(1 - \nu), & \text{if } Y = 0.
\end{cases}
\end{equation}
For $Y=0$ the probability that it won't rain is expressed by $1-\nu$. For $Y>0$ the probability that it will rain $\nu$ is multiplied with the left-truncated at zero density function. Contrary to the censored model, here it is only the parameter $\nu$ controlling the point mass while $\mu$ and $\sigma$ only influence the distribution of the positive observations.
\end{itemize}

@@

In the log-likelihood function in \ref{eq:hurdle} properties of the
logarithmic function can be applied to split the term for $Y>0$ 
into two additive terms separating the parameters $\mu$ and $\sigma$
from the additional parameter $\nu$. Therefore, $\nu$ can be modeled 
by a separate model in the GAMLSS framework. This approach is also
chosen for the distributional forest model in this application study.

@@

@@

\section{Results}
In Figure~\ref{fig:lattice} each panel belongs to one of the 15 observation stations. The three distributions (left-censored at zero Gaussian distribution = \textit{gaussian}, hurdle model based on the Gaussian distribution = \textit{hgaussian}, left-censored at zero logistic distribution = \textit{logistic}) are compared based on the resulting CRPS values for each of the applied methods that are represented by different colours and symbols. Overall it can be stated that the forest model performs almost equally well with all three distributions. Therefore it seems to be very robust against missspecifications of the distribution.
For the two GAMLSS models the censored Gaussian and the hurdle model perform similarly at most of the station while applying the censored logistic distribution leads to an increase in CRPS, i.e. to a decline in performance, especially for the prespecified GAMLSS.
For the EMOS model exchanging the distribution does not have a strong effect on its performance, similar to the forest model. In particular, for the two censored distributions the results are almost equal, only for the hurdle model the CRPS values vary slightly, however in both directions, depending on the station.

When applying a hurdle model for the stations Oetz and Innvervillgraten the prespecified GAMLSS computationally failed to build/fit a model. Therefore this method is not represented in Figure~\ref{fig:lattice} for these two cases.

In stations like Ladis-Neuegg, Matrei in Osttirol or Oetz all methods seem to be very robust and stable over varying distributions. The strongest influence of an exchange of the distribution can be seen for example at stations Jungholz and St.Johann im Walde for the prespecified GAMLSS.

In Figure\ref{fig:box_crpss_gaussian} the tested distributions are compared for each method based on CRPS skill score with the censored Gaussian distribution as reference represented by the horizontal line at zero. The hurdle model leads to a slight improvement for the forest model and the prespecified GAMLSS. Following the median this improvement is a bit stronger for the boosted GAMLSS, however, with a higher deviation as represented by the larger size of the box, similarly as for the EMOS, however with a negative influence indicated by the negative median. Choosing a logistic distribution instead of a Gaussian clearly reduces the strength of the two GAMLSS models, while it has almost no influence on the forest and the EMOS models.

\begin{figure}[h]
\setkeys{Gin}{width=0.99\textwidth}
<<results_lattice, fig=TRUE, echo=FALSE, results=hide, height=10, width=8>>=
library("lattice")
##### 
# HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)

load("Rain_distributions.rda")
means <- results$means[results$means$method != "disttree",]
means$method <- factor(means$method, 
                       levels = c("distforest",
                                  "gamlss",
                                  "gamboostLSS",
                                  "EMOS"),
                       labels = c("Distributional forest",
                                  "Prespecified GAMLSS",
                                  "Boosted GAMLSS",
                                  "EMOS"))

@@

xyplot(crps ~ distribution | station, groups = ~ method, 
       data = means, auto.key = TRUE, 
       type = "o", lwd = 2, lty = 1, layout = c(3,6))
@
\caption{\label{fig:lattice}CRPS values of all 4 methods, once applied with each of the three distributions for each of the 15 observation stations.}
\end{figure}

\begin{figure}[h]
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_gaussian, fig=TRUE, echo=FALSE, height=3, width=7>>=
  # crps skill score by distribution (reference: gaussian)
  means$crps_ss_dist <- means$crps
  means$crps_ss_dist[means$distribution == "gaussian"] <- 
    1 - means$crps[means$distribution == "gaussian"] / means$crps[means$distribution == "gaussian"]
  means$crps_ss_dist[means$distribution == "logistic"] <- 
    1 - means$crps[means$distribution == "logistic"] / means$crps[means$distribution == "gaussian"]
  means$crps_ss_dist[means$distribution == "hgaussian"] <- 
    1 - means$crps[means$distribution == "hgaussian"] / means$crps[means$distribution == "gaussian"]

@@

  par(mfrow = c(1,4))
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "Distributional forest"), main = "Distributional forest",
          ylim = c(-0.12,0.07), las = 2, ylab = "CRPS skill score")
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "Prespecified GAMLSS"), main = "Prespecified GAMLSS",
          ylim = c(-0.12,0.07), las = 2)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "Boosted GAMLSS"), main = "Boosted GAMLSS",
          ylim = c(-0.12,0.07), las = 2)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "EMOS"), main = "EMOS",
          ylim = c(-0.12,0.07), las = 2)
  abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_gaussian}(Mean) CRPS skill score values of each of the 15 observation stations, separated by method and distribution. Gaussian is chosen as reference distribution.}
\end{figure}

\begin{figure}[h]
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_EMOS, fig=TRUE, echo=FALSE, height=3, width=7>>=
  # crps skill score by method (reference: EMOS)
  means$crps_ss_method <- means$crps
  means$crps_ss_method[means$method == "Distributional forest"] <- 
    1 - means$crps[means$method == "Distributional forest"] / means$crps[means$method == "EMOS"]
  means$crps_ss_method[means$method == "Prespecified GAMLSS"] <- 
    1 - means$crps[means$method == "Prespecified GAMLSS"] / means$crps[means$method == "EMOS"]
  means$crps_ss_method[means$method == "Boosted GAMLSS"] <- 
    1 - means$crps[means$method == "Boosted GAMLSS"] / means$crps[means$method == "EMOS"]
  means$crps_ss_method[means$method == "EMOS"] <- 
    1 - means$crps[means$method == "EMOS"] / means$crps[means$method == "EMOS"]

@@

  par(mfrow = c(1,3))
  boxplot(crps_ss_method~method, data = means_sel, 
          subset = (distribution == "gaussian"), main = "gaussian",
          ylim = c(-0.12,0.23), axes = FALSE, ylab = "CRPS skill score")
  axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 2)
  axis(1, 0:4, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS", ""), las=2)
  axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_method~method, data = means_sel, 
          subset = (distribution == "logistic"), main = "logistic",
          ylim = c(-0.12,0.23), axes = FALSE)
  axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 2)
  axis(1, 0:4, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS", ""), las=2)
  axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_method~method, data = means_sel, 
          subset = (distribution == "hgaussian"), main = "hgaussian",
          ylim = c(-0.12,0.23), axes = FALSE)
  axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 2)
  axis(1, 0:4, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS", ""), las=2)
  axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
  abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_EMOS}(Mean) CRPS skill score values of each of the 15 observation stations, separated by method and distribution. EMOS is chosen as reference model.}
\end{figure}

\begin{figure}[h]
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crps, fig=TRUE, echo=FALSE, height=3, width=7>>=

>>>>

\documentclass[nojss, shortnames]{jss}

@@

\Abstract{In the application discussed in the paper ''Distributional 
Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain''
the underlying distribution is assumed to be a left-censored at zero Gaussian distribution.
In this supplement the exact same evaluations as in the paper are performed employing
other distribution assumptions. In particular, a left-censored at zero logistic distribution 
and a hurdle-model are applied and their results are compared to those of the left-censored 
at zero Gaussian distribution presented in the paper.}

\Keywords{random forests, GAMLSS, hurdle model, logistic, Gaussian}

@@

Torsten Hothorn\\
Universit\"at Z\"urich \\
Institut f\"ur Epidemiologie, Biostatistik und Pr\"avention \\
Hirschengraben 84\\
CH-8001 Z\"urich, Switzerland \\
E-mail: \email{Torsten.Hothorn@R-project.org}\\
URL: \url{http://user.math.uzh.ch/hothorn/}\\
}

@@

\section{Introduction}
This application study is supplementary material to the application presented
in Section~3 of the paper.
There, the presented distributional forest model is employed to obtain 
probabilistic precipitation forecasts in a mountainous region assuming that the 
response variable (the power-transformed daily precipitation sum) follows a 
left-censored Gaussian distribution with censoring point~$0$.

@@

To compare the performance of the distributional forest to other commonly used
methods the exact same models as in the paper are evaluated on the same data set 
and applying the same settings while only the distribution is exchanged. Further 
details on the models can be found in Section~3.2. of the paper. In particular, 
Table~2 gives an overview over the models and their specifications. As a brief 
reminder the approaches are summarized below.

@@

\item \emph{Distributional forest:} All predictor variables are considered for 
learning a forest where subsampling is employed for each tree. Parameters are 
estimated by adaptive local likelihood based on forest weights.

@@

The distribution applied in the paper is based on a Gaussian distribution 
for which the probability density function $f$ and the distribution 
function $F$ are defined by
%
\begin{align}
f_{\mathrm{Gaussian}}(Y; \mu, \sigma) &= 
\frac{1}{\sigma} \cdot \phi\left(\frac{Y-\mu}{\sigma}\right) \\
F_{\mathrm{Gaussian}}(Y; \mu, \sigma) &=
\Phi\left(\frac{Y-\mu}{\sigma}\right)
\end{align}
where $\phi$ and $\Phi$ are the probability density function and the 
distribution function of the standard normal distribution $\mathcal{N}(0,1)$.

Alternatively one could also choose a distribution with heavier tails 
such as a logistic distribution with corresponding probability density 
function and distribution function
%
\begin{align}
f_{\mathrm{logistic}}(Y; \mu, \sigma) &= 
\frac{\exp{\left(-\frac{Y - \mu}{\sigma}\right)}}
{\sigma \cdot \left(1 + \exp{\left(-\frac{Y - \mu}{\sigma}\right)}\right)^2} \\
F_{\mathrm{logistic}}(Y; \mu, \sigma) &=
\frac{1}{1+ \exp{\left(-\frac{Y - \mu}{\sigma}\right)}}.
\end{align}
%
Both distributions have to be adapted in our application to capture that
precipitation is non-negative with a point mass at zero (= no rain).
Two common strategies for this are to employ either \emph{left-censoring at zero}
or a \emph{hurdle} (or two-part) approach, combining a binary (zero vs.\ greater) and
a truncated (positive) part \citep[see][for an overview]{Winkelmann:2006}.
\begin{itemize}
\item Censored distribution \citep{Winkelmann:2006}: 
%In general, censoring data usually leads 
%to a point mass at the censoring point which has to be considered in 
%the modeling process. Therefore, censored distributions are applied as 
%they are able to capture this point mass. 
%In precipitation data, observations can not be negative, thus, left-censoring
%it at zero does not change the data. However, there is a naturally given
%point mass at zero (days without rain) making a censored distribution an 
%appropriate choice to model precipitation data.
For a chosen distribution with probability density function $f$ and
distribution function $F$ the log-likelihood function of the corresponding 
censored distribution is
%
\begin{equation}
\ell_{\mathrm{cens}}(\mu, \sigma; Y) = 
\begin{cases}
\log\left\{f(Y; \mu, \sigma) \right\}, & \text{if } Y > 0 \\[0.2cm]
\log\left\{F(0; \mu, \sigma) \right\}, & \text{if } Y = 0
\end{cases}
\end{equation}
%
as seen in Equation~2.1.1. of the paper for a Gaussian distribution.
For positive values of $Y$ the censored version of the density function
equals the original non-censored version. For $Y=0$ the censored density
function takes the values of the probability mass lying left of the 
censoring point. Therefore, the parameters $\mu$ and $\sigma$ control
both parts, the point mass and the positive part of the distribution.

\item Hurdle model / Two-part model (\cite{Cragg:1971}, \cite{Mullahy:1986}): 
An additional parameter $\nu$ describing the probability that there is 
any precipitation at all, i.e. $\nu = \mathbb{P}(Y>0)$, is included in the model.
The three-parametric log-likelihood function for the full hurdle model 
with parameter vector $\bm{\theta} = (\mu, \sigma, \nu)$ is
\begin{equation} \label{eq:hurdle}
\ell_{\mathrm{hurdle}}(\mu, \sigma, \nu; Y) = 
\begin{cases}
\log\left\{\nu \cdot \frac{f(Y; \mu, \sigma)}{1 - F(0; \mu, \sigma)} \right\}, & \text{if } Y > 0 \\[0.2cm]
\log(1 - \nu), & \text{if } Y = 0.
\end{cases}
\end{equation}
For $Y=0$ the probability that it won't rain is expressed by $1-\nu$. 
For $Y>0$ the probability that it will rain $\nu$ is multiplied with the 
left-truncated at zero density function. Contrary to the censored model, 
here it is only the parameter $\nu$ controlling the point mass while $\mu$ 
and $\sigma$ only influence the distribution of the positive observations.

@@

In the log-likelihood function in Equation~\ref{eq:hurdle} properties of the
logarithmic function can be applied to split the term for $Y>0$ 
into two additive terms separating the parameters $\mu$ and $\sigma$
from the additional parameter $\nu$. Therefore, $\nu$ can be modeled 
by a separate model in the GAMLSS framework. The same approach is
chosen for the distributional forest model in this application study.

@@

Selected stations: Axams, Lech, Zuers, See im Paznaun, Jungholz, Ladis-Neuegg, 
Oetz, Ochsengarten-Obergut, Ginzling, Rotholz, Walchsee, Koessen, 
Innervillgraten, Matrei in Osttirol, St.Johann im Walde.

@@

\section{Results}
In Figure~\ref{fig:lattice} each panel belongs to one of the 15 observation stations. 
The three distributions \textit{gaussian} (left-censored at zero Gaussian distribution), 
\textit{hgaussian} (hurdle model based on the Gaussian distribution) and 
\textit{logistic} (left-censored at zero logistic distribution) are compared based 
on the resulting CRPS values for each of the four applied methods that are represented 
by different colors and symbols. 
Overall, the forest model performs almost equally well for all three distributions. 
Therefore it seems to be very robust against missspecifications of the distribution.
For the two GAMLSS models the censored Gaussian and the hurdle model perform similarly 
at most of the stations while applying the censored logistic distribution leads to an 
increase in CRPS, i.e. to a decline in performance, especially for the prespecified 
GAMLSS (for example at stations Zuers and Jungholz).
For the EMOS model exchanging the distribution does not have a strong effect on its 
performance, similar to the forest model.
All in all, exchanging the underlying distribution does not lead to clear 
improvements of any of the methods, especially the censored Gaussian and the 
hurdle model achieve almost the same results. However, the robustness against 
missspecification of the distribution varies slightly over stations. For example 
at station Matrei in Osttirol all methods seem to be very stable while stronger 
influences can be seen at station Jungholz.

\textit{Note:} When applying a hurdle model for the stations Oetz and 
Innvervillgraten the prespecified GAMLSS computationally failed to build/fit 
a model. Therefore this method is not represented in the corresponding panels 
of Figure~\ref{fig:lattice} for these two cases.

In Figure~\ref{fig:box_crpss_gaussian} the tested distributions are compared 
for each method based on CRPS skill score with the censored Gaussian distribution 
as reference represented by the horizontal line at zero. Again, the forest model 
and the EMOS show only slight variations (on average not even 1\%) over different 
distributions while the two GAMLSS models perform even or slightly better applying 
a hurdle model and loose strength when employing a logistic distribution compared 
to a Gaussian distribution.

\begin{figure}
\setkeys{Gin}{width=0.99\textwidth}
<<results_lattice, fig=TRUE, echo=FALSE, results=hide, height=10, width=8>>=
library("lattice")
##### 
# HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)

load("Rain_distributions.rda")
stations <- c("Axams", "Lech", "Zuers", "See im Paznaun", "Jungholz", 
              "Ladis-Neuegg", "Oetz", "Ochsengarten-Obergut",
              "Ginzling", "Rotholz", "Walchsee", "Koessen", 
              "Innervillgraten", "Matrei in Osttirol", 
              "St.Johann im Walde")

means <- results$means[(results$means$method != "disttree") & 
                         results$means$station %in% stations, ]
rownames(means) <- c(1:NROW(means))
means$station <- factor(means$station, levels = stations)
means$method <- factor(means$method, 
                       levels = c("distforest",
                                  "gamlss",
                                  "gamboostLSS",
                                  "EMOS"),
                       labels = c("Distributional forest",
                                  "Prespecified GAMLSS",
                                  "Boosted GAMLSS",
                                  "EMOS"))

@@

xyplot(crps ~ distribution | station, groups = ~ method, 
       data = means, auto.key = TRUE, 
       type = "o", lwd = 2, lty = 1,  
       drop.unused.levels = TRUE, layout = c(3,5),
       as.table = TRUE, ylab = "CRPS", xlab = "Distribution")
@
\caption{\label{fig:lattice}CRPS values of all four methods, once applied with 
each of the three distributions for each of the 15 observation stations.}
\end{figure}

\begin{figure}
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_gaussian, fig=TRUE, echo=FALSE, height=3, width=7>>=
# crps skill score by distribution (reference: gaussian)
means$crps_ss_dist <- means$crps
means$crps_ss_dist[means$distribution == "gaussian"] <- 
  1 - means$crps[means$distribution == "gaussian"] / means$crps[means$distribution == "gaussian"]
means$crps_ss_dist[means$distribution == "logistic"] <- 
  1 - means$crps[means$distribution == "logistic"] / means$crps[means$distribution == "gaussian"]
means$crps_ss_dist[means$distribution == "hgaussian"] <- 
  1 - means$crps[means$distribution == "hgaussian"] / means$crps[means$distribution == "gaussian"]

@@

# CRPS skill score for distributional forest
s_df <- matrix(ncol = 2, nrow = 15)
s_df[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_df[,2] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "logistic", "crps_ss_dist"]

# CRPS skill score for prespecified GAMLSS
s_g <- matrix(ncol = 2, nrow = 15)
s_g[,1] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_g[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "logistic", "crps_ss_dist"]

# CRPS skill score for boosted GAMLSS
s_gb <- matrix(ncol = 2, nrow = 15)
s_gb[,1] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_gb[,2] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "logistic", "crps_ss_dist"]

# CRPS skill score for EMOS
s_emos <- matrix(ncol = 2, nrow = 15)
s_emos[,1] <- means_sel[means_sel$method == "EMOS" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_emos[,2] <- means_sel[means_sel$method == "EMOS" & means_sel$distribution == "logistic", "crps_ss_dist"]

colnames(s_df) <- colnames(s_g) <- colnames(s_gb) <- colnames(s_emos) <- c("hgaussian", "logistic")

# boxplots with matplot
par(mfrow = c(1,4), mar = c(4,4,3,0.7))
matplot(t(s_df[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "Distributional forest",
        xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_df, add = TRUE, col = "transparent")
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Distributional forest"), main = "Distributional forest",
#        ylim = c(-0.12,0.07), las = 2, ylab = "CRPS skill score")
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,2.9,3,1.8))
matplot(t(s_g[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Prespecified GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_g, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Prespecified GAMLSS"), main = "Prespecified GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,1.8,3,2.9))
matplot(t(s_gb[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Boosted GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_gb, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Boosted GAMLSS"), main = "Boosted GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,0.7,3,4))
matplot(t(s_emos[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "EMOS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_emos, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "EMOS"), main = "EMOS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_gaussian} CRPS skill score values of each of 
the 15 observation stations, separated by method and distribution. Left-censored 
at zero Gaussian is chosen as reference distribution.}
\end{figure}

Figures~\ref{fig:box_crpss_EMOS} and \ref{fig:box_crps} compare the methods separately 
for each distribution, once based on CRPS skill score values with reference model EMOS 
and once based on CRPS values. Both confirm the conclusion that the censored Gaussian 
model and the hurdle model lead to almost the same results for all methods while only 
the logistic model shows small variations.

Altogether it can be stated that the left-censored at zero Gaussian distribution seems 
to be an appropriate choice while the other two distributions, especially the hurdle model, 
are also reasonable alternatives. However, they do not lead to an overall improvement in 
performance for this application.

\begin{figure}
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_EMOS, fig=TRUE, echo=FALSE, height=3, width=7>>=
# crps skill score by method (reference: EMOS)
means$crps_ss_method <- means$crps
means$crps_ss_method[means$method == "Distributional forest"] <- 
  1 - means$crps[means$method == "Distributional forest"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "Prespecified GAMLSS"] <- 
  1 - means$crps[means$method == "Prespecified GAMLSS"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "Boosted GAMLSS"] <- 
  1 - means$crps[means$method == "Boosted GAMLSS"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "EMOS"] <- 
  1 - means$crps[means$method == "EMOS"] / means$crps[means$method == "EMOS"]

@@

# CRPS skill score for gaussian
s_gaussian <- matrix(ncol = 3, nrow = 15)
s_gaussian[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "gaussian", "crps_ss_method"]
s_gaussian[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "gaussian", "crps_ss_method"]
s_gaussian[,3] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "gaussian", "crps_ss_method"]

# CRPS skill score for boosted hgaussian
s_hgaussian <- matrix(ncol = 3, nrow = 15)
s_hgaussian[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "hgaussian", "crps_ss_method"]
s_hgaussian[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "hgaussian", "crps_ss_method"]
s_hgaussian[,3] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "hgaussian", "crps_ss_method"]

# CRPS skill score for prespecified logistic
s_logistic <- matrix(ncol = 3, nrow = 15)
s_logistic[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "logistic", "crps_ss_method"]
s_logistic[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "logistic", "crps_ss_method"]
s_logistic[,3] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "logistic", "crps_ss_method"]

colnames(s_gaussian) <- colnames(s_hgaussian) <- colnames(s_logistic) <- c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS")

par(mfrow = c(1,3), mar = c(5,4,3,0))
matplot(t(s_gaussian[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "gaussian", ylim = c(-0.12,0.23), 
        xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 3.5))
boxplot(s_gaussian, add = TRUE, col = "transparent", axes = FALSE)
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_method~method, data = means_sel, 
#        subset = (distribution == "gaussian"), main = "gaussian",
#        ylim = c(-0.12,0.23), axes = FALSE, ylab = "CRPS skill score")
axis(1, 0:4, c("",
               "Distr.
                 forest",
               "Presp.
                 GAMLSS",
               "Boosted
                 GAMLSS", ""), las=2)
axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 0)
axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
